Efficient Compilation of High Level Python Numerical Programs with Pythran

### 2014-08-22 doc/papers/cgo2014/paper.tex
Improve Stackoverflow part

>>>>

One of the first steps required when designing an optimizing compiler is to
gather enough test cases, or benchmarks, in order to drive the optimization
process toward realistic examples. To achieve this goal, we have gathered a few
exemple from existing compiler sources, but these may have been modified to
overcome some compiler limitations. A less biased source of benchmarks should
be independant from the compiler community and focused on matching user needs.
To achieve that goal, we collected synthetic scientifc kernels from the
stackoverflow website.

StackOverflow~\footnote{\url{http://stackoverflow.com/}} is a question and
answer site for programmers. Any registered user can submit questions and other
users may answer. A voting system is then used to sort the answers. Said
otherwise, the website provides a database of question and answers on
programming topics, including Python.

A query using the terms \emph{numpy} and \emph{slow} yields many results, many
of which use the pattern ``Q: my code is slow'' followed by an upvoted answer
``A: rewrite it that way'', where the original code performs explicit looping
over an array, and the rewritten code uses the relevant combination of numpy
function.

For instance, question \url{http://stackoverflow.com/questions/7741878}
proposes to use an explicit loop to iteratively compute the $L^2$ norm of each
row of a matrix, as illustrated in Listing~\ref{lst:l2norm-loopy}, and one of
the proposed answer is to use the more compact version presented in
Listing~\ref{lst:l2norm}. This answer no longer uses any explicit loops and
roughly achieves a $times 4$ speedup over the loop version.

\begin{lstlisting}[language=python,caption={Per row version of $L^2$ norm with loop in numpy.}, float, label={lst:l2norm-loopy}]
def slow(x):
    r = np.empty(x.shape[0])
    for i in xrange(x.shape[0]):
        r[i] = np.sum(np.abs(x[i])**2)
    return r
\end{lstlisting}

\begin{lstlisting}[language=python,caption={Per row version of $L^2$ norm without loop in numpy.}, float, label={lst:l2norm}]
import numpy as np
def l2norm(x):
    return np.sum(np.abs(x)**2, axis=1)
\end{lstlisting}

Our guess is that new Numpy user are more likely to write loops and code
similar to Listing~\ref{lst:l2norm-loopy}, but as their knowledge of the numpy
API and of the good programming practice grows, they get to write code similar
to Listing~\ref{lst:l2norm}. A direct consequence is that a compiler should
focus on optimizing high-level constructs rather loops, both to encourage good
practice with respect to Numpy code, and because that's the kind of code Numpy
user ultimately write.
